---
title: "TMS WM Analysis"
output: html_document
date: "2025-1-29"
---

## Packages

```{r}
# import packages
library(lme4)     # mixed models
library(dplyr)    # dataframe manipulation
library(readr)    # reading in the data
library(emmeans)  # likelihood ratio testing
library(purrr)    # for map()
library(lattice)  # for xyplot()
library(TAF)      # setting up output file directory structure during model fitting
library(moments)  # skewness for selecting between additive and multiplicative models of response times
library(stringr)


#If you get an error, you likely need to install the packages using the "install.package('insert-package-name-here')" command
```

## Notebook configurable settings

```{r}
# name the directories to read data from and write output to
input_version = 9
output_version = 'programming'
program_type = 'short'

```

make an overall output directory (model summaries, stats, etc. will all go here)

```{r}
#setwd("..")
outputdir = sprintf("%s/../candidate-models/regression-models-%s",getwd(), output_version)
```

```{r}

mkdir(outputdir)
```

## Loading Data

```{r}
#setwd("..")
input_file = sprintf("%s/../preprocessed-data/full-data-version-%d/programming-full-data.csv", getwd(), input_version)
data <- read.csv(input_file)

#Transforming the session number such that it is treated as a character and not a value
data <- transform( 
  data, session = as.character(session)) 

#Transforming correctness into binaries of 1 or 0
data$correctness <- as.integer(ifelse(data$accuracy == 'Pass', 1, 0))

```

Removing irrelevant data from dataframe

```{r}
#remove rows that have no values - nothing actually happens here. Everyone saw the stimuli
data <- data[data$duration > 1, ] 


#Removing the one participants that didn't listen to instructions (2 SD below the mean in N-back):
participants_to_remove <- c("REDACTED_20")
data <- subset(data, !(participant_id %in% participants_to_remove))
```

Looking at FITB and the passed ones only - Make sure there are 533 rows

```{r}

data <- data[data$problem_type == program_type, ]


number_of_rows <- nrow(data)
print(number_of_rows) #number of rows

```

Make sure you have n=27 participants

```{r}
#Make sure you have n=27 participants

id <- unique(data$participant_id)
id #all unique IDs
length(id) #number of participants
```

Mean correctness

```{r}
#Make sure you have n=27 participants

mean_value <- mean((data$correctness))
std_value <- sd((data$correctness))
print(mean_value)
print(std_value)
```

Mean corectness is 0.69 with a standard deviation of 0.46

Table of problems views by participants. Some participants saw only 7-8 problems per session while some saw 11-12 per session.

```{r}
count_table <- table(data$TMS_Type, data$participant_id)

print(count_table)
```

## Select independent variables

1.  `participant_id` (person) : random intercept
    -   People are drawn from an underlying distribution; we aren't interested in learning coefficients specific to the members of our sample, but rather in what they tell us about that underlying population. Random effects tell us about the distribution rather than the elements observed.

    -   This variable is used to model person-specific attributes, such as a "skill," translating to overall faster or slower response times, that can be different for each person
2.  `problem_id` (question prompt) : random intercept
    -   The underlying distribution of stimuli is a little less clear than that of participants. This variable can be thought of as modeling the dispersion of question-specific difficulties.
3.  `TMS_Type` (TMS treatment condition, i.e. brain region stimulated) : fixed & random intercepts
    -   Based on previous literature, we hypothesize that the brain region stimulated may have an effect that is systematic (i.e. very consistent between people, topics, etc.), quantitatively heterogeneous (i.e. directionally consistent, but widely varying in magnitude), or even a qualitatively heterogeneous (i.e. varying in direction, possibly in conjunction with magnitude). Therefore, we assess the goodness of fit for candidate models containing `TMS_Type` as a fixed effect, random effect, both, or neither, in order to test this hypothesis.
4.  `session`and `coding_section_no` (chronological numbering of the session, per participant) : fixed intercept
    -   This is used to account for learning/training effect, i.e. participants getting better at test-taking under our specific test instrument over the course of the study. For example, they may become more familiar with the question structure, or become better at recalling old material as they do more practice questions. We expect this to be systematic (uniformly decreasing response time), so we model it as a fixed effect.

## Specify candidate predictor structures for mixed model formulas

```{r}
fixed_effects = c(
  # no fixed effects
  "", 
  
  # no interactions
  "TMS_Type", 
  "session", 
  "coding_section_no", 
  "problem_order", 
  "problem_set", 
  
  "pain",
  "physical_symptoms",
  "discomfort", 
  "sensations",
  
  "TMS_Type + session",
  "TMS_Type + coding_section_no",
  "TMS_Type + problem_order",
  "TMS_Type + problem_set",
  "session + coding_section_no",
  "session + problem_order",
  "session + problem_set",
  "problem_set + coding_section_no",
  "problem_set + problem_order",
  "coding_section_no + problem_order",
  
  "TMS_Type + pain",
  "TMS_Type + physical_symptoms",
  "TMS_Type + discomfort",
  "TMS_Type + sensations",
  "session + pain",
  "session + physical_symptoms",
  "session + discomfort",
  "session + sensations",
  "coding_section_no + pain",
  "coding_section_no + physical_symptoms",
  "coding_section_no + discomfort",
  "coding_section_no + sensations",
  "problem_order + pain",
  "problem_order + physical_symptoms",
  "problem_order + discomfort",
  "problem_order + sensations",
  "problem_set + pain",
  "problem_set + physical_symptoms",
  "problem_set + discomfort",
  "problem_set + sensations",
  
  "TMS_Type + session + coding_section_no + problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",
  
  # 2-way interactions
  "TMS_Type * session",
  "TMS_Type * coding_section_no",
  "TMS_Type * problem_order",
  "TMS_Type * problem_set",
  "session * coding_section_no",
  "session * problem_order",
  "session * problem_set",
  "problem_set * coding_section_no",
  "problem_set * problem_order",
  "coding_section_no * problem_order",

  "TMS_Type * pain",
  "TMS_Type * physical_symptoms",
  "TMS_Type * discomfort",
  "TMS_Type * sensations", 
  "session * pain",
  "session * physical_symptoms",
  "session * discomfort",
  "session * sensations",
  "coding_section_no * pain",
  "coding_section_no * physical_symptoms",
  "coding_section_no * discomfort",
  "coding_section_no * sensations",
  "problem_order * pain",
  "problem_order * physical_symptoms",
  "problem_order * discomfort",
  "problem_order * sensations",
  "problem_set * pain",
  "problem_set * physical_symptoms",
  "problem_set * discomfort",
  "problem_set * sensations",
  
  # 1- and 2-way interactions
  "TMS_Type * session + problem_set + coding_section_no + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session + coding_section_no + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * coding_section_no + session + problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session + problem_set + coding_section_no + pain + physical_symptoms + discomfort + sensations",
  "session * problem_set + TMS_Type + coding_section_no + problem_order + pain + physical_symptoms + discomfort + sensations",
  "session * coding_section_no + TMS_Type + problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "session * problem_order + TMS_Type + problem_set + coding_section_no + pain + physical_symptoms + discomfort + sensations",
  "problem_set * coding_section_no + TMS_Type + session + problem_order + pain + physical_symptoms + discomfort + sensations",
  "problem_set * problem_order + TMS_Type + session + coding_section_no + pain + physical_symptoms + discomfort + sensations",
  "coding_section_no * problem_order + TMS_Type + session + problem_set + pain + physical_symptoms + discomfort + sensations",
  
  "TMS_Type * pain + session + problem_set + coding_section_no + problem_order",
  "TMS_Type * physical_symptoms + session + problem_set + coding_section_no + problem_order",
  "TMS_Type * discomfort + session + problem_set + coding_section_no + problem_order",
  "TMS_Type * sensations + session + problem_set + coding_section_no + problem_order", 
  
  # multiple 2-way interactions
  "TMS_Type * session + problem_set * coding_section_no + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * session + problem_set * problem_order + coding_section_no + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * session + coding_section_no * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session * coding_section_no + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session * problem_order + coding_section_no + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + coding_section_no * problem_order + session + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * coding_section_no + session * problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * coding_section_no + session * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * coding_section_no + problem_set * problem_order + session + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session * problem_set + coding_section_no + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session * coding_section_no + problem_set + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + problem_set * coding_section_no + session + pain + physical_symptoms + discomfort + sensations",
  "session * problem_set + coding_section_no * problem_order + TMS_Type + pain + physical_symptoms + discomfort + sensations",
  "session * coding_section_no + problem_set * problem_order + TMS_Type + pain + physical_symptoms + discomfort + sensations",
  "session * problem_order + problem_set * coding_section_no + TMS_Type + pain + physical_symptoms + discomfort + sensations",
  
  # 3-way interactions
  "TMS_Type * session * problem_set + coding_section_no + problem_order + pain + physical_symptoms + discomfort + sensations", 
  "TMS_Type * session * coding_section_no + problem_set + problem_order + pain + physical_symptoms + discomfort + sensations", 
  "TMS_Type * session * problem_order + problem_set + coding_section_no + pain + physical_symptoms + discomfort + sensations", 
  "TMS_Type * problem_set * coding_section_no + problem_order + session + pain + physical_symptoms + discomfort + sensations", 
  "TMS_Type * problem_set * problem_order + coding_section_no + session + pain + physical_symptoms + discomfort + sensations", 
  "TMS_Type * coding_section_no * problem_order + session + problem_set + pain + physical_symptoms + discomfort + sensations", 
  "session * problem_set * coding_section_no + problem_order + TMS_Type + pain + physical_symptoms + discomfort + sensations", 
  "session * problem_set * problem_order + coding_section_no + TMS_Type + pain + physical_symptoms + discomfort + sensations", 
  "session * coding_section_no * problem_order + TMS_Type + problem_set + pain + physical_symptoms + discomfort + sensations", 
  "problem_set * coding_section_no * problem_order + TMS_Type + session + pain + physical_symptoms + discomfort + sensations"
)
```

```{r}
random_effects = c(
  #"",                               # (1)
  "participant_id", 
  "problem_id", 
  #"domain",                         # (2)
  #"condition",                      # (2)
  #"id:stimulus",                    # (3)
  #"participant_id:problem_type",
  "participant_id:TMS_Type", 
  #"stimulus:domain",                # (4)
  "problem_id:TMS_Type" #, 
  #"domain:condition",               # (2)
  #"id:stimulus:domain",             # (3), (4)
  #"id:stimulus:condition",          # (3)
  #"participant_id:problem_type:TMS_Type"
  #"stimulus:domain:condition",      # (2), (4)
  #"id:stimulus:domain:condition"    # (2), (3), (4)
)
```

In this first cell, we enumerate random effect structures by taking all combinations of the random effects specified.

```{r}
random_effects = map(random_effects, function (s) sprintf("(1|%s)", s))

random_exprs = c()
N_random = length(random_effects)
for (k in 1:N_random) {
  iii = combn(1:N_random, k, simplify=FALSE)
  for (ii in iii) {
    expr = paste(random_effects[ii], collapse = " + ")
    random_exprs = append(random_exprs, expr)
  }
}
```

Now, we combine the response variable, fixed effects, and random effects together to generate our candidate model structures.

```{r}
logistic_formulae = c()
for (fe in fixed_effects) {
  for (re in random_exprs) {
    f = ""
    if (fe == "") {
      f = sprintf("correctness ~ %s", re)
    } else {
      f = sprintf("correctness ~ %s + %s", fe, re)
    }
    logistic_formulae = append(logistic_formulae, f)
  }
}

print(length(logistic_formulae))
```

Filter out invalid model specifications

```{r}
# generate all the boolean index vectors

# formulas with TMS fixed effect
has_TMS <- grepl('TMS_Type ', logistic_formulae, fixed = TRUE)

# formulas with participant random effect
has_participant <- grepl('(1|participant_id)', logistic_formulae, fixed = TRUE)

# formulas with problem random effect
has_problem <- grepl('(1|problem_id)', logistic_formulae, fixed = TRUE)

# formulas with TMS x participant random effect
has_TMS_participant <- grepl('(1|participant_id:TMS_Type)', logistic_formulae, fixed = TRUE)

# formulas with TMS x problem random effect
has_TMS_problem <- grepl('(1|problem_id:TMS_Type)', logistic_formulae, fixed = TRUE)

# remove them

invalid <- (has_TMS_participant & (!has_TMS | !has_participant)) | (has_TMS_problem & (!has_TMS | !has_problem))

invalid_logistic_formulae <- logistic_formulae[invalid]
logistic_formulae <- logistic_formulae[!invalid]

print(length(logistic_formulae))
```

## Model fitting (\~2 minutes)

Here, we fit all the models. Of note, we use MLE rather than REML for compatibility with the AIC model selection strategy.

We also store BIC (an alternative model selection criterion) in addition to AIC here. The two have different theoretical properties, both of which may be useful. Both describe the log likelihood of a fitted model, with a penalty for model complexity. The penalty in BIC is higher than that for AIC, so optimal-BIC models tend to be smaller (with fewer parameters) than optimal-AIC models. We have done our analysis relying on AIC, but we provide BIC-based model selection here in case the reader is interested.

```{r}
# make an output directory for the output of all the candidate linear models
correctness_outputdir = 
  sprintf("%s/FITB_v9-2_acc_model_summaries", outputdir)
mkdir(correctness_outputdir)
```

```{r fit_all_models, cache=TRUE}
if (TRUE) {
  # make an output file for the AIC and BIC of each model
  metricsfile = sprintf("%s/metrics.csv", correctness_outputdir)
  file.create(metricsfile)
  
  sink(file = metricsfile,
       append = FALSE,
       type = "output",
       split = FALSE
       )
  
  cat("formula,AIC,BIC\n")
  
  sink()
  
  for (f in logistic_formulae) {
    # make an output directory for the output of this specific candidate model
    f_outputdir = sprintf("%s/%s", correctness_outputdir, gsub(" ", "", f, fixed = TRUE))
    mkdir(f_outputdir)
    
    # make the individual output file
    summaryfile = sprintf("%s/summary.txt", f_outputdir)
    
    file.create(summaryfile)
    
    # print a delimiter for console output
    cat(sprintf("\n---------------------------------------------\n%s\n", f))
    
    # fit the model
    #m = try(glmer(f, family=binomial, data=data))
    
    result <- tryCatch({
      # Code that may produce an error
      m = try(glmer(f, family=binomial, data=data))
      # send the model summary to the summary file
      sink(file = summaryfile,
         append = FALSE,
         type = "output",
         split = FALSE
         )
    
      print(summary(m))
    
      sink()
      
      # send the AIC and BIC to the metrics file
      metrics = sprintf("%s,%f,%f\n", f, AIC(m), BIC(m))
    
      sink(file = metricsfile,
         append = TRUE,
         type = "output",
         split = FALSE)
    
      cat(metrics)
    
      sink()
      
      }, error = function(e) {
      # Handle the error
      cat("An error occurred:", conditionMessage(e), "\n")
      NA
      })
    
  }
}
```

## Model selection

```{r}
# read in the AIC/BIC data
#fin = sprintf("%s/FITB_MWP_babytest_model_summaries/metrics.csv", outputdir) #from original submission
fin = sprintf("%s/FITB_v9-2_acc_model_summaries/metrics.csv", outputdir)
correctness_models_da = read_csv(fin)
dim(correctness_models_da)
```

```{r}
head(correctness_models_da[order(correctness_models_da$AIC),], n = 10)
```

```{r}
head(correctness_models_da[order(correctness_models_da$BIC),], n = 10)
```

Our selection process pulls `correctness ~ TMS_Type + coding_section_no + (1|problem_id) + (1|participant_id:TMS_Type)` as the best model to represent the data by AIC (and 2nd-best by BIC).

```{r}
fA <- correctness_models_da[order(correctness_models_da$AIC),]$formula[[1]]
cat(fA)
```

BIC removes exactly the `TMS_Type` terms.

```{r}
fB <- correctness_models_da[order(correctness_models_da$BIC),]$formula[[1]]
cat(fB)
```

```{r}
m0 <- glmer(fA, family=binomial, data=data)
summary(m0)
```

The fixed effect is not significant. Next we check if `(1 | participant_id:TMS_Type)` has statistical significance.

### Check whether to include the factor

```{r}
m1 <- glmer(fB, family=binomial, data=data)
summary(m1)
```

```{r}
anova(m0, m1)
```

This means including the `TMS_Type` variable is significant. Again though, it is *not* the marginal effect:

```{r}
# both with interaction term
m0_tmp <- glmer(
  correctness ~ TMS_Type + coding_section_no + (1 | participant_id) + (1 | problem_id), 
  family=binomial, 
  data=data
)
m1_tmp <- glmer(
  correctness ~ coding_section_no + (1 | participant_id) + (1 | problem_id), 
  family=binomial, 
  data=data
)
anova(m0_tmp, m1_tmp)
```

```{r}
# both without interaction term
m0_tmp <- glmer(
  correctness ~ TMS_Type + coding_section_no + (1 | participant_id) + (1 | problem_id) + (1 | participant_id:TMS_Type), 
  family=binomial, 
  data=data
)
m1_tmp <- glmer(
  correctness ~ coding_section_no + (1 | participant_id) + (1 | problem_id) + (1 | participant_id:TMS_Type), 
  family=binomial, 
  data=data
)
anova(m0_tmp, m1_tmp)
```

### Check if random effect is statistically significant

```{r}
m2 <- glmer(
  correctness ~ TMS_Type + coding_section_no + (1 | participant_id) + (1 | problem_id), 
  family=binomial, 
  data=data
)
anova(m0, m2)
```

This is actually a stronger p-value than the 'omnibus' test removing both terms. So the random interaction intercept is significant (although the confint below should be reported, rather than this p-value).

## Confidence Intervals and PSVE

### Point estimate confint

```{r}
summary(m0)
```

```{r profile_pp0, cache=TRUE}
pp0 = profile(m0)
```

```{r}
confint(pp0)
```

```{r}
xyplot(pp0)
```

$\sigma_1$ excludes zero, so this further confirms the likelihood ratio test above.

### Percentage of Structural Variance Explained

```{r}
get_psve = function(m, effect, save=FALSE, fout, effect_order) {
  # m is a glmer object so residual variance isn't well-defined, hence 'structural'
  # The lower confidence bound returned by bootMer is within standard tolerance of zero, 
  # even though the lower confidence bound of the variance estimate alone is far from 
  # zero. To understand why, I want to examine the variance estimates for each intercept 
  # and simulation, but because dataframes are copy-on-modify in R, I have to write them 
  # to a file to store them.
  random_effects = as.data.frame(VarCorr(m)) %>% 
    rename(Groups = grp, Name = var1, Variance = vcov, Std.Dev. = sdcor) %>% 
    select(-var2)
  
  num = (random_effects %>% filter(Groups == effect))$Variance
  denom = sum(random_effects$Variance)
  
  ## side effect
  if (save) {
    if (missing(fout) || missing(effect_order)) {
      warning(sprintf(
        'get_psve(m, effect=%s, save=TRUE) failed to save: fout or effect_order not provided', 
        effect))
    } else {
      vv <- sapply(effect_order, function(eff) {
        (random_effects %>% filter(Groups == eff))$Variance
      }) # this is to ensure that they appear in the same order every time
      sink(
        file = fout,
        append = TRUE,
        type = "output",
        split = FALSE
      )
      cat(paste0(vv, collapse=','), '\n')
      sink()
    }
  }
  ##
  
  return (c(num / denom))
}

get_psve(m0, "participant_id:TMS_Type")
```

```{r}
#tmp <- as.data.frame(VarCorr(m0))
#tmp$grp
VarCorr(m0)
```

### PSVE confint

```{r bootmer_b0, cache=TRUE}
ordered_factors <- as.data.frame(VarCorr(m0))$grp
nsim <- 10000
# 10k sims took about 2.5 hrs

for (nsim in c(8000, 9000, 11000, 12000)) {
  
}
psve_log <- sprintf("%s/psve_bootstrap_%i.csv", correctness_outputdir, nsim)
file.create(psve_log)
cat(psve_log, "\n")
# write the header
sink(
  file = psve_log,
  append = FALSE,
  type = "output",
  split = FALSE
)
cat(paste0(ordered_factors, collapse=","), "\n")
sink()

b0 = bootMer(
  m0, 
  FUN = function (m) {return(get_psve(m, "participant_id:TMS_Type", save=TRUE, psve_log, ordered_factors))}, 
  nsim = nsim,
  .progress = "txt"
)
```

```{r}

```

```{r}
quantile(b0$t, c(0.025, 0.975))
```

```{r}
bootstrap_sims <- read.csv(psve_log)
bootstrap_sims$sum <- rowSums(bootstrap_sims)
bootstrap_sims$psve <- bootstrap_sims$participant_id.TMS_Type / bootstrap_sims$sum
bootstrap_sims <- bootstrap_sims %>% arrange(psve) 
bootstrap_sims %>% head(50)
```

```{r}
quantile(bootstrap_sims$participant_id.TMS_Type, c(0.025, 0.975))
quantile(bootstrap_sims$psve, c(0.025, 0.975))
```

```{r}
bottom_percentile <- 3
bottom_percentile_idx <- (nsim * bottom_percentile / 100)
```

```{r}
# Distribution
hist(bootstrap_sims$psve, breaks=20, right=FALSE)
abline(v = quantile(b0$t, 0.025), col = "red", lwd = 2, lty = 2) # 2.5 percentile
abline(v = get_psve(m0, "participant_id:TMS_Type"), col = "black", lwd = 2, lty = 2) # estimate
abline(v = quantile(b0$t, 0.975), col = "red", lwd = 2, lty = 2) # 97.5 percentile

# Low-PSVE simulations
plot(bootstrap_sims[1:bottom_percentile_idx,]$psve, 
     type="o", 
     main=sprintf('Simulations in bottom %d%% PSVE', 
                  bottom_percentile), 
     xlab='Simulation #', 
     ylab='PSVE')
abline(v = 0.025*nsim, col = "red", lwd = 2, lty = 2) # 2.5 percentile

# All simulations
plot(bootstrap_sims$psve, 
     type="l", 
     main='PSVE for all simulations', 
     xlab='Simulation #', 
     ylab='PSVE')
abline(v = 0.025*nsim, col = "red", lwd = 2, lty = 2) # 2.5 percentile
abline(v = nsim/2, col = "blue", lwd = 2, lty = 2) # median
abline(v = 0.975*nsim, col = "red", lwd = 2, lty = 2) # 97.5 percentile

```

```{r}
# Num and denom in low-PSVE simulations
plot(bootstrap_sims[1:bottom_percentile_idx,]$participant_id.TMS_Type, 
     type="o", 
     main=sprintf('TMS:participant variance (numerator) of simulations in bottom %d%% PSVE', bottom_percentile), 
     xlab='Simulation #', 
     ylab='Value')
abline(v = 0.025*nsim, col = "red", lwd = 2, lty = 2) # 2.5 percentile

plot(bootstrap_sims[1:bottom_percentile_idx,]$sum, 
     type="o", 
     main=sprintf('Sum of variances (denominator) of simulations in bottom %d%% PSVE', bottom_percentile), 
     xlab='Simulation #', 
     ylab='Value')
abline(v = 0.025*nsim, col = "red", lwd = 2, lty = 2) # 2.5 percentile

# Component variances in low-PSVE simulations
all_y_values <- c(
  bootstrap_sims[1:bottom_percentile_idx,]$participant_id.TMS_Type, 
  bootstrap_sims[1:bottom_percentile_idx,]$participant_id, 
  bootstrap_sims[1:bottom_percentile_idx,]$problem_id)
y_range <- range(all_y_values)

plot(bootstrap_sims[1:bottom_percentile_idx,]$participant_id.TMS_Type, 
     type="o", 
     ylim=y_range, 
     main=sprintf('Variance components of simulations in bottom %d%% PSVE', bottom_percentile), 
     xlab='Simulation #', 
     ylab='Value')
lines(bootstrap_sims[1:bottom_percentile_idx,]$participant_id, type="o")
lines(bootstrap_sims[1:bottom_percentile_idx,]$problem_id, type="o")
abline(v = 0.025*nsim, col = "red", lwd = 2, lty = 2) # 2.5 percentile
```

```{r}
# call this 'structural variance ratio' or 'structural PVE' - will check the name
# go back and compute structural PVE for paper 1
# intra-class correlation, g-ratio
# x1. get df of 1000 sims x 3 var estimates
# x2. make PVE column and sort by that from smallest
# x3. look at head and see if numerator got small or denom got big (in first 2.5%)
# ?? profit
```

### Other thing

```{r}
VarCorr(m0)
```

Manual Calculation of CI We needed to do this since the profiler function went to error.

```{r}
# Check the summary of the model
model_summary <- summary(m0)

# Extract the random effects variance estimates
random_effects_var <- as.data.frame(VarCorr(m0))

# View the variance components
#print(random_effects_var)

# Calculate confidence intervals for the random effects
# Can calculate the CI based on a normal approximation:
z_value <- 1.96  # For 95% confidence intervals

# Create a data frame to hold CI results
ci_results <- data.frame(
  Group = random_effects_var$grp,
  Var_Estimate = random_effects_var$vcov,
  CI_Lower = random_effects_var$vcov - z_value * sqrt(random_effects_var$vcov),
  CI_Upper = random_effects_var$vcov + z_value * sqrt(random_effects_var$vcov)
)

# Print the confidence intervals for random effects
print(ci_results)
```

'participant_id:TMS_Type' effect does not exclude zero which means it is not significant

We may also be interested in assessing the differences in accuracy outcome fom control to treatment

```{r}
#How often did people get an answer right in the Control group vs Treatment group?
df_summary <- data %>%
  group_by(TMS_Type) %>%
  summarise(
    mean_correct = mean(correctness, na.rm = TRUE),
    sd_correct = sd(correctness, na.rm = TRUE),
    count = n()
  )

print(df_summary)

```

Since correctness is typically represented as 0 (incorrect) or 1 (correct) in binary data. Mean correctness is 0.71 in control and 0.67 in WM with SD of 0.4 in both an SD of 0.4 is quite large. This suggests that some participants answered almost all questions correctly (near 1.0), while others answered very few correctly (near 0.3 or lower). The variation might be due to individual differences, task difficulty, or randomness in responses.

## Test other model candidates

```{r}
best_AIC <- correctness_models_da[order(correctness_models_da$AIC),]$AIC[[1]]
best_AIC

AIC_candidates <- correctness_models_da[correctness_models_da$AIC < best_AIC + 3,]
AIC_candidates <- AIC_candidates[order(AIC_candidates$AIC),]
AIC_candidates
```

```{r}
# formulas with TMS fixed effect
has_TMS <- grepl('TMS_Type ', AIC_candidates$formula, fixed = TRUE)

add_fixed_TMS <- paste0(AIC_candidates$formula, ' + TMS_Type')
AIC_candidates$fixed_TMS_full <- ifelse(
  has_TMS, AIC_candidates$formula, add_fixed_TMS)

AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$formula, fixed('TMS_Type + '))
AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$fixed_TMS_reduced, fixed('TMS_Type * '))

AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$fixed_TMS_reduced, fixed(' + (1|participant_id:TMS_Type)'))
AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$fixed_TMS_reduced, fixed(' + (1|problem_id:TMS_Type)'))
# unsure whether (1|participant_id:TMS_Type) should be removed in the reduced version
```

```{r}
# debugging
for (f in AIC_candidates$fixed_TMS_full){
  cat(f, '\n')
}
cat('\n\n')
for (f in AIC_candidates$fixed_TMS_reduced){
  cat(f, '\n')
}
```

```{r}
AIC_candidates$fixed_TMS_pval <- -1.0
for (i in seq_len(nrow(AIC_candidates))) {
  m_full <- glmer(
    AIC_candidates$fixed_TMS_full[i],
    data = data,
    family = binomial
  )
  m_reduced <- glmer(
    AIC_candidates$fixed_TMS_reduced[i],
    data = data,
    family = binomial
  )
  a <- anova(m_full, m_reduced)
  p_val <- a$"Pr(>Chisq)"[2]
  #print(p_val)
  AIC_candidates$fixed_TMS_pval[i] <- p_val
}

AIC_candidates[AIC_candidates$fixed_TMS_pval < 0.05,]
```

This turns out to be more involved, and there are only 2 besides the original one so I'll do it manually.

```{r}
# 1. formula with fixed interaction
# a) test random interaction - significant
f_full_tmp <- AIC_candidates$formula[[2]]
f_red_tmp <- 'correctness ~ TMS_Type * coding_section_no + (1|participant_id) + (1|problem_id)'

m_full_tmp <- glmer(
  f_full_tmp,
  data = data,
  family = binomial
)
m_reduced_tmp <- glmer(
  f_red_tmp,
  data = data,
  family = binomial
)
anova(m_full_tmp, m_reduced_tmp)

cat('\n\n')

# b) test fixed interaction - not significant
f_red_tmp <- 'correctness ~ TMS_Type + coding_section_no + (1|participant_id) + (1|problem_id) + (1 | participant_id:TMS_Type)'
m_reduced_tmp <- glmer(
  f_red_tmp,
  data = data,
  family = binomial
)
anova(m_full_tmp, m_reduced_tmp)

# c) test marginal
# same as original tests (best-AIC model)
```

```{r}
# 2. formula with additional random effect - not significant
f_full_tmp <- AIC_candidates$formula[[3]]
f_red_tmp <- 'correctness ~ TMS_Type + coding_section_no + (1|participant_id) + (1|problem_id) + (1 | participant_id:TMS_Type)'

m_full_tmp <- glmer(
  f_full_tmp,
  data = data,
  family = binomial
)
m_reduced_tmp <- glmer(
  f_red_tmp,
  data = data,
  family = binomial
)
anova(m_full_tmp, m_reduced_tmp)
```
