---
title: "TMS WM Analysis"
output: html_document
date: "2025-1-29"
---

## 0. Packages

```{r}
# import packages
library(lme4) # mixed models
library(dplyr) # dataframe manipulation
library(readr) # reading in the data
library(emmeans) # likelihood ratio testing
library(purrr) # for map()
library(lattice) # for xyplot()
library(TAF) # setting up output file directory structure during model fitting
library(moments) # skewness for selecting between additive and multiplicative models of response times
library(stringr)


# If you get an error, you likely need to install the packages using the "install.package('insert-package-name-here')" command
```

## 1. Notebook configurable settings

```{r}
# name the directories to read data from and write output to
input_version <- 9
output_version <- "n-back"
nback_version <- "1-back"
```

```{r}
# setwd("..")
outputdir <- sprintf("%s/../candidate-models/regression-models-%s", getwd(), output_version)
```

```{r}
mkdir(outputdir)
```

## Loading Data

```{r}
# setwd("..")
input_file <- sprintf(
  "%s/../preprocessed-data/full-data-version-%d/nback-full-data.csv",
  getwd(), input_version
)
data <- read.csv(input_file)



# Transforming the session number such that it is treated as a character and not a value
# data <- transform(
# data, session = as.character(session))


data <- transform(
  data,
  correct_answer = as.integer(as.logical(correct_answer))
)

head(data)
```

### Removing irrelevant data from dataframe

```{r Nans}
# remove rows that have no values - nothing actually happens here. Everyone saw the stimuli
data <- data[data$response.time != 0, ]


# Removing the one participants that didn't listen to instructions:
participants_to_remove <- c("REDACTED_20")
data <- subset(data, !(participant_id %in% participants_to_remove))
```

### Making sure we are only looking at one type of n-back at a time

```{r}
data <- data[data$nback_type == nback_version, ] # 1-back , 2-back , or 3-back
```

### Determining what is the mean response.time

```{r}
mean_value <- mean((data$response.time))
print(mean_value)
```

### Checking number of participants is n=27

```{r}
id <- unique(data$participant_id)
id # all unique IDs
length(id) # number of participants
```

### Checking distribution of of observartions per factor load in n-back

```{r}
count_table <- table(data$TMS_Type, data$nback_type)

# Print the result
print(count_table)
```

```{r}
fixed_effects <- c(
  # no fixed effects
  "",

  # no interactions
  "TMS_Type",
  "session",
  "problem_order",
  "problem_set",
  "pain",
  "physical_symptoms",
  "discomfort",
  "sensations",

  "TMS_Type + session",
  "TMS_Type + problem_order",
  "TMS_Type + problem_set",
  "session + problem_order",
  "session + problem_set",
  "problem_set + problem_order",

  "TMS_Type + pain",
  "TMS_Type + physical_symptoms",
  "TMS_Type + discomfort",
  "TMS_Type + sensations",
  "session + pain",
  "session + physical_symptoms",
  "session + discomfort",
  "session + sensations",
  "problem_order + pain",
  "problem_order + physical_symptoms",
  "problem_order + discomfort",
  "problem_order + sensations",
  "problem_set + pain",
  "problem_set + physical_symptoms",
  "problem_set + discomfort",
  "problem_set + sensations",

  "TMS_Type + session + problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",

  # 2-way interactions
  "TMS_Type * session",
  "TMS_Type * problem_order",
  "TMS_Type * problem_set",
  "session * problem_order",
  "session * problem_set",
  "problem_set * problem_order",

  "TMS_Type * pain",
  "TMS_Type * physical_symptoms",
  "TMS_Type * discomfort",
  "TMS_Type * sensations",
  "session * pain",
  "session * physical_symptoms",
  "session * discomfort",
  "session * sensations",
  "problem_order * pain",
  "problem_order * physical_symptoms",
  "problem_order * discomfort",
  "problem_order * sensations",
  "problem_set * pain",
  "problem_set * physical_symptoms",
  "problem_set * discomfort",
  "problem_set * sensations",

  # 1- and 2-way interactions
  "TMS_Type * session + problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session + problem_set + pain + physical_symptoms + discomfort + sensations",
  "session * problem_set + TMS_Type + problem_order + pain + physical_symptoms + discomfort + sensations",
  "session * problem_order + TMS_Type + problem_set + pain + physical_symptoms + discomfort + sensations",
  "problem_set * problem_order + TMS_Type + session + pain + physical_symptoms + discomfort + sensations",

  "TMS_Type * pain + session + problem_set + problem_order",
  "TMS_Type * physical_symptoms + session + problem_set + problem_order",
  "TMS_Type * discomfort + session + problem_set + problem_order",
  "TMS_Type * sensations + session + problem_set + problem_order",

  # multiple 2-way interactions
  "TMS_Type * session + problem_set * problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session * problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session * problem_set + pain + physical_symptoms + discomfort + sensations",

  # 3-way interactions
  "TMS_Type * session * problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * session * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set * problem_order + session + pain + physical_symptoms + discomfort + sensations",
  "session * problem_set * problem_order + TMS_Type + pain + physical_symptoms + discomfort + sensations"
)
```

```{r}
random_effects <- c(
  "participant_id",
  "shown_letter",
  # other than adding this term, all is the same as in the programming analyses
  # (it couldn't be included in the programming analyses because there would
  # only be one observation per group)
  "participant_id:shown_letter",
  "participant_id:TMS_Type",
  "shown_letter:TMS_Type"
)
```

In this first cell, we enumerate random effect structures by taking all combinations of the random effects specified in Section 5.

```{r}
random_effects <- map(random_effects, function(s) sprintf("(1|%s)", s))

random_exprs <- c()
N_random <- length(random_effects)
for (k in 1:N_random) {
  iii <- combn(1:N_random, k, simplify = FALSE)
  for (ii in iii) {
    expr <- paste(random_effects[ii], collapse = " + ")
    random_exprs <- append(random_exprs, expr)
  }
}
```

Now, we combine the response variable, fixed effects, and random effects together to generate our candidate model structures.

```{r}
logistic_formulae <- c()
for (fe in fixed_effects) {
  for (re in random_exprs) {
    f <- ""
    if (fe == "") {
      f <- sprintf("(correct_answer) ~ %s", re)
    } else {
      f <- sprintf("(correct_answer) ~ %s + %s", fe, re)
    }
    logistic_formulae <- append(logistic_formulae, f)
  }
}

print(length(logistic_formulae))
```

Finally, filter out invalid model specifications.

```{r}
# generate all the boolean index vectors

# formulas with TMS fixed effect
has_TMS <- grepl("TMS_Type ", logistic_formulae, fixed = TRUE)

# formulas with participant random effect
has_participant <- grepl("(1|participant_id)", logistic_formulae, fixed = TRUE)

# formulas with problem random effect
has_problem <- grepl("(1|shown_letter)", logistic_formulae, fixed = TRUE)

# formulas with TMS x participant random effect
has_TMS_participant <- grepl(
  "(1|participant_id:TMS_Type)", logistic_formulae, fixed = TRUE
)

# formulas with TMS x problem random effect
has_TMS_problem <- grepl(
  "(1|shown_letter:TMS_Type)", logistic_formulae, fixed = TRUE
)

# formulas with participant x problem random effect
has_participant_problem <- grepl(
  "(1|participant_id:shown_letter)", logistic_formulae, fixed = TRUE
)

# remove them

invalid <- (
  has_TMS_participant & (!has_TMS | !has_participant)) |
  (has_TMS_problem & (!has_TMS | !has_problem)) |
  (has_participant_problem & (!has_participant | !has_problem)
)

invalid_logistic_formulae <- logistic_formulae[invalid]
logistic_formulae <- logistic_formulae[!invalid]

print(length(logistic_formulae))
```

## Model fitting (\~2 minutes)

Here, we fit all the models. Of note, we use MLE rather than REML for compatibility with the AIC model selection strategy.

We also store BIC (an alternative model selection criterion) in addition to AIC here. The two have different theoretical properties, both of which may be useful. Both describe the log likelihood of a fitted model, with a penalty for model complexity. The penalty in BIC is higher than that for AIC, so optimal-BIC models tend to be smaller (with fewer parameters) than optimal-AIC models. We have done our analysis relying on AIC, but we provide BIC-based model selection here in case the reader is interested.

```{r fit_all_models, cache=TRUE}
if (TRUE) {
  # make an output directory for the output of all the candidate linear models
  correctness_outputdir <-
    sprintf("%s/oneback_v9_acc_model_summaries", outputdir)
  mkdir(correctness_outputdir)

  # make an output file for the AIC and BIC of each model
  metricsfile <- sprintf("%s/metrics.csv", correctness_outputdir)
  file.create(metricsfile)

  sink(
    file = metricsfile,
    append = FALSE,
    type = "output",
    split = FALSE
  )

  cat("formula,AIC,BIC\n")

  sink()

  for (f in logistic_formulae) {
    # make an output directory for the output of this specific candidate model
    f_outputdir <- sprintf(
      "%s/%s", correctness_outputdir, gsub(" ", "", f, fixed = TRUE)
    )
    mkdir(f_outputdir)

    # make the individual output file
    summaryfile <- sprintf("%s/summary.txt", f_outputdir)

    file.create(summaryfile)

    # print a delimiter for console output
    cat(sprintf("\n---------------------------------------------\n%s\n", f))

    # fit the model
    # m = try(glmer(f, family=binomial, data=data))

    result <- tryCatch(
      {
        # Code that may produce an error
        m <- try(glmer(f, family = binomial, data = data))
        # send the model summary to the summary file
        sink(
          file = summaryfile,
          append = FALSE,
          type = "output",
          split = FALSE
        )

        print(summary(m))

        sink()

        # send the AIC and BIC to the metrics file
        metrics <- sprintf("%s,%f,%f\n", f, AIC(m), BIC(m))

        sink(
          file = metricsfile,
          append = TRUE,
          type = "output",
          split = FALSE
        )

        cat(metrics)

        sink()
      },
      error = function(e) {
        # Handle the error
        cat("An error occurred:", conditionMessage(e), "\n")
        NA
      }
    )
  }
}
```

## Model selection

```{r}
# read in the AIC/BIC data
fin <- sprintf("%s/oneback_v9_acc_model_summaries/metrics.csv", outputdir)
correctness_models_da <- read_csv(fin)
```

```{r}
head(correctness_models_da[order(correctness_models_da$AIC), ], n = 10)
```

```{r}
head(correctness_models_da[order(correctness_models_da$BIC), ], n = 10)
```

## Optimal AIC

```{r}
fA <- correctness_models_da[order(correctness_models_da$AIC), ]$formula[[1]]
cat(fA)
```

```{r}
fB <- correctness_models_da[order(correctness_models_da$BIC), ]$formula[[1]]
cat(fB)
```

Neither selected model includes a TMS_Type effect.

```{r}
m0 <- lmer(fA, data = data, REML = FALSE)
summary(m0)
```

## Testing other model candidates

For completeness, we try the likelihood ratio test with all low-AIC model formulas. If there was a positive result here, we would have to correct for multiple tests, but there isn't.

```{r}
best_AIC <- correctness_models_da[order(correctness_models_da$AIC), ]$AIC[[1]]
best_AIC

AIC_candidates <- correctness_models_da[correctness_models_da$AIC < best_AIC + 3, ]
AIC_candidates <- AIC_candidates[order(AIC_candidates$AIC), ]
AIC_candidates
```

```{r}
# formulas with TMS fixed effect
has_TMS <- grepl('TMS_Type ', AIC_candidates$formula, fixed = TRUE)

add_fixed_TMS <- paste0(AIC_candidates$formula, ' + TMS_Type')
AIC_candidates$fixed_TMS_full <- ifelse(
  has_TMS, AIC_candidates$formula, add_fixed_TMS)

AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$formula, fixed('TMS_Type + '))
AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$fixed_TMS_reduced, fixed('TMS_Type * '))

AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$fixed_TMS_reduced, fixed(' + (1|participant_id:TMS_Type)'))
AIC_candidates$fixed_TMS_reduced <- str_remove_all(
  AIC_candidates$fixed_TMS_reduced, fixed(' + (1|problem_id:TMS_Type)'))
# unsure whether (1|participant_id:TMS_Type) should be removed in the reduced version
```

```{r}
AIC_candidates$fixed_TMS_pval <- -1.0
for (i in seq_len(nrow(AIC_candidates))) {
  m_full <- glmer(
    AIC_candidates$fixed_TMS_full[i],
    data = data,
    family = binomial
  )
  m_reduced <- glmer(
    AIC_candidates$fixed_TMS_reduced[i],
    data = data,
    family = binomial
  )
  a <- anova(m_full, m_reduced)
  p_val <- a$"Pr(>Chisq)"[2]
  #print(p_val)
  AIC_candidates$fixed_TMS_pval[i] <- p_val
}

AIC_candidates[AIC_candidates$fixed_TMS_pval < 0.05,]
```


