---
title: "1-back Reaction Time Analysis"
output: html_document
date: "2025-1-29"
---

## Packages

```{r}
# import packages
library(lme4) # mixed models
library(dplyr) # dataframe manipulation
library(readr) # reading in the data
library(emmeans) # likelihood ratio testing
library(purrr) # for map()
library(lattice) # for xyplot()
library(TAF) # setting up output file directory structure during model fitting
library(moments) # skewness for selecting between additive and multiplicative models of response times
library(filelock)
library(lmerTest)

# If you get an error, you likely need to install the packages using the "install.package('insert-package-name-here')" command
```

## Notebook configurable settings

```{r}
# name the directories to read data from and write output to
input_version <- 9
output_version <- "n-back"
nback_version <- "1-back"
```

make an overall output directory (model summaries, stats, etc. will all go here)

```{r}
# setwd("..")
outputdir <- sprintf(
  "%s/../candidate-models/regression-models-%s", getwd(), output_version
)
```

```{r}
mkdir(outputdir)
```

## Loading Data

```{r}
# setwd("..")
input_file <- sprintf("%s/../preprocessed-data/full-data-version-%d/nback-full-data.csv", getwd(), input_version)
data <- read.csv(input_file)


# Transforming the session number such that it is treated as a character and not a value
data <- transform(
  data,
  session = as.character(session)
)

head(data)
```

Removing irrelevant data from dataframe

```{r}
# remove rows that have no values - nothing actually happens here. Everyone saw the stimuli
# data <- data[data$response.time != 0, ]


# Removing the one participants that didn't listen to instructions:
participants_to_remove <- c("REDACTED_20")
data <- subset(data, !(participant_id %in% participants_to_remove))
```

Making sure we are only looking at one type of n-back at a time and only instance of correct response

```{r}
data <- data[data$correct_answer == "True", ] # only counting the onces they got correct


data <- data[data$nback_type == nback_version, ] # 1-back , 2-back , or 3-back
```

Determining what is the mean response.time

```{r}
mean_value <- mean((data$response.time))
print(mean_value)
```

Centralizing the response.time so that it is more normally distributed. One-back was not generally normally distributed as seen in the bottom graph so transformation is required

```{r}
print(skewness(data$response.time))
print(skewness(sqrt(data$response.time)))
print(skewness(log(data$response.time))) # best
```

```{r}
hist(log(data$response.time))
```

Checking number of participants is n=27

```{r}
id <- unique(data$participant_id)
id # all unique IDs
length(id) # number of participants
```

Checking distribution of of observations per factor load in n-back

```{r}
count_table <- table(data$TMS_Type, data$nback_type)

# Print the result
print(count_table)
```

## Select independent variables

1.  `participant_id` (person) : random intercept
    -   People are drawn from an underlying distribution; we aren't interested in learning coefficients specific to the members of our sample, but rather in what they tell us about that underlying population. Random effects tell us about the distribution rather than the elements observed.

    -   This variable is used to model person-specific attributes, such as a "skill," translating to overall faster or slower response times, that can be different for each person
2.  `shown_letter` (letter presented) : random intercept
    -   The underlying distribution of stimuli is a little less clear than that of participants: specifically since stimuli in N-back are just a random fixed set of letters. However, prior work on N-back have found that some letters may be easier to remember than others. We control for that here.
3.  `TMS_Type` (TMS treatment condition, i.e. brain region stimulated) : fixed & random intercepts
    -   Based on previous literature, we hypothesize that the brain region stimulated may have an effect that is systematic (i.e. very consistent between people, topics, etc.), quantitatively heterogeneous (i.e. directionally consistent, but widely varying in magnitude), or even a qualitatively heterogeneous (i.e. varying in direction, possibly in conjunction with magnitude). Therefore, we assess the goodness of fit for candidate models containing `TMS_Type` as a fixed effect, random effect, both, or neither, in order to test this hypothesis.
4.  `session` (chronological numbering of the session, per participant) : fixed intercept
    -   This is used to account for learning/training effect, i.e. participants getting better at test-taking under our specific test instrument over the course of the study. For example, they may become more familiar with the question structure, or become better at recalling old material as they do more practice questions. We expect this to be systematic (uniformly decreasing response time), so we model it as a fixed effect.

## Specify candidate predictor structures for mixed model formulas

```{r}
fixed_effects <- c(
  # no fixed effects
  "",

  # no interactions
  "TMS_Type",
  "session",
  "problem_order",
  "problem_set",
  "pain",
  "physical_symptoms",
  "discomfort",
  "sensations",

  "TMS_Type + session",
  "TMS_Type + problem_order",
  "TMS_Type + problem_set",
  "session + problem_order",
  "session + problem_set",
  "problem_set + problem_order",

  "TMS_Type + pain",
  "TMS_Type + physical_symptoms",
  "TMS_Type + discomfort",
  "TMS_Type + sensations",
  "session + pain",
  "session + physical_symptoms",
  "session + discomfort",
  "session + sensations",
  "problem_order + pain",
  "problem_order + physical_symptoms",
  "problem_order + discomfort",
  "problem_order + sensations",
  "problem_set + pain",
  "problem_set + physical_symptoms",
  "problem_set + discomfort",
  "problem_set + sensations",

  "TMS_Type + session + problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",

  # 2-way interactions
  "TMS_Type * session",
  "TMS_Type * problem_order",
  "TMS_Type * problem_set",
  "session * problem_order",
  "session * problem_set",
  "problem_set * problem_order",

  "TMS_Type * pain",
  "TMS_Type * physical_symptoms",
  "TMS_Type * discomfort",
  "TMS_Type * sensations",
  "session * pain",
  "session * physical_symptoms",
  "session * discomfort",
  "session * sensations",
  "problem_order * pain",
  "problem_order * physical_symptoms",
  "problem_order * discomfort",
  "problem_order * sensations",
  "problem_set * pain",
  "problem_set * physical_symptoms",
  "problem_set * discomfort",
  "problem_set * sensations",

  # 1- and 2-way interactions
  "TMS_Type * session + problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session + problem_set + pain + physical_symptoms + discomfort + sensations",
  "session * problem_set + TMS_Type + problem_order + pain + physical_symptoms + discomfort + sensations",
  "session * problem_order + TMS_Type + problem_set + pain + physical_symptoms + discomfort + sensations",
  "problem_set * problem_order + TMS_Type + session + pain + physical_symptoms + discomfort + sensations",

  "TMS_Type * pain + session + problem_set + problem_order",
  "TMS_Type * physical_symptoms + session + problem_set + problem_order",
  "TMS_Type * discomfort + session + problem_set + problem_order",
  "TMS_Type * sensations + session + problem_set + problem_order",

  # multiple 2-way interactions
  "TMS_Type * session + problem_set * problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set + session * problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_order + session * problem_set + pain + physical_symptoms + discomfort + sensations",

  # 3-way interactions
  "TMS_Type * session * problem_set + problem_order + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * session * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations",
  "TMS_Type * problem_set * problem_order + session + pain + physical_symptoms + discomfort + sensations",
  "session * problem_set * problem_order + TMS_Type + pain + physical_symptoms + discomfort + sensations"
)
```

```{r}
random_effects <- c(
  "participant_id",
  "shown_letter",
  # other than adding this term, all is the same as in the programming analyses
  # (it couldn't be included in the programming analyses because there would
  # only be one observation per group)
  "participant_id:shown_letter",
  "participant_id:TMS_Type",
  "shown_letter:TMS_Type"
)
```

In this first cell, we enumerate random effect structures by taking all combinations of the random effects specified in Section 5.

```{r}
random_effects <- map(random_effects, function(s) sprintf("(1|%s)", s))

random_exprs <- c()
N_random <- length(random_effects)
for (k in 1:N_random) {
  iii <- combn(1:N_random, k, simplify = FALSE)
  for (ii in iii) {
    expr <- paste(random_effects[ii], collapse = " + ")
    random_exprs <- append(random_exprs, expr)
  }
}
```

Now, we combine the response variable, fixed effects, and random effects together to generate our candidate model structures.

```{r}
loglinear_formulae <- c()
for (fe in fixed_effects) {
  for (re in random_exprs) {
    f <- ""
    if (fe == "") {
      f <- sprintf("log(response.time) ~ %s", re)
    } else {
      f <- sprintf("log(response.time) ~ %s + %s", fe, re)
    }
    loglinear_formulae <- append(loglinear_formulae, f)
  }
}

print(length(loglinear_formulae))
```

### Model fitting (\~2 minutes)

Here, we fit all the models. Of note, we use MLE rather than REML for compatibility with the AIC model selection strategy.

We also store BIC (an alternative model selection criterion) in addition to AIC here. The two have different theoretical properties, both of which may be useful. Both describe the log likelihood of a fitted model, with a penalty for model complexity. The penalty in BIC is higher than that for AIC, so optimal-BIC models tend to be smaller (with fewer parameters) than optimal-AIC models. We have done our analysis relying on AIC, but we provide BIC-based model selection here in case the reader is interested.

```{r fit_all_models, cache=TRUE}
if (TRUE) {
  # make an output directory for the output of all the candidate linear models
  logtime_outputdir <-
    sprintf("%s/oneback_v9_log_time_model_summaries", outputdir)
  mkdir(logtime_outputdir)

  # make an output file for the AIC and BIC of each model
  metricsfile <- sprintf("%s/metrics.csv", logtime_outputdir)
  file.create(metricsfile)

  sink(
    file = metricsfile,
    append = FALSE,
    type = "output",
    split = FALSE
  )

  cat("formula,AIC,BIC\n")

  sink()

  for (f in loglinear_formulae) {
    # make an output directory for the output of this specific candidate model
    f_outputdir <- sprintf("%s/%s", logtime_outputdir, gsub(" ", "", f, fixed = TRUE))
    mkdir(f_outputdir)

    # make the individual output file
    summaryfile <- sprintf("%s/summary.txt", f_outputdir)

    file.create(summaryfile)

    # print a delimiter for console output
    cat(sprintf("\n---------------------------------------------\n%s\n", f))

    # fit the model
    m <- lmer(f, data = data, REML = FALSE)

    # send the model summary to the summary file
    sink(
      file = summaryfile,
      append = FALSE,
      type = "output",
      split = FALSE
    )

    print(summary(m))

    sink()

    # send the AIC and BIC to the metrics file
    metrics <- sprintf("%s,%f,%f\n", f, AIC(m), BIC(m))

    sink(
      file = metricsfile,
      append = TRUE,
      type = "output",
      split = FALSE
    )

    cat(metrics)

    sink()
  }
}
```

### Model selection

```{r}
# read in the AIC/BIC data
fin <- sprintf("%s/oneback_v9_log_time_model_summaries/metrics.csv", outputdir)
log_time_models_da <- read_csv(fin)
```

```{r}
head(log_time_models_da[order(log_time_models_da$AIC), ], n = 10)
```

```{r}
head(log_time_models_da[order(log_time_models_da$BIC), ], n = 10)
```

### Optimal AIC

```{r}
fA <- log_time_models_da[order(log_time_models_da$AIC), ]$formula[[1]]
cat(fA)
```

```{r}
fB <- log_time_models_da[order(log_time_models_da$BIC), ]$formula[[1]]
cat(fB)
```

The AIC favors the model `log(response.time) ~ TMS_Type * session * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations + (1|participant_id) + (1|participant_id:TMS_Type)` however BIC favors a different model `log(response.time) ~ TMS_Type * problem_order + (1|participant_id) + (1|participant_id:TMS_Type)`.

```{r}
m0 <- lmer(fA, data = data, REML = FALSE)
summary(m0)
```

```{r}
m1 <- lmer(fB, data = data, REML = FALSE)
summary(m1)
```

## Which model is better: AIC- or BIC-selected?
This step is not really part of our analysis procedure, because in general the two model specifications are not always nested so not always comparable by likelihood ratio test. However, they happen to be nested in this case, so we might as well double-check.

If Pr(\>Chisq) \< 0.05 → m0 is significantly better than m1 ✅ Keep m0, since the extra parameter improves model fit. If Pr(\>Chisq) \> 0.05 → No significant difference ❌ Prefer m1 (simpler model), since m0 doesn’t improve fit enough.

```{r}
anova(m0, m1)
```

The larger model `log(response.time) ~ TMS_Type * session * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations + (1|participant_id) + (1|participant_id:TMS_Type)` is a significantly better fit.

## Omnibus test TMS effects

Specify the nested model with all TMS effect terms removed.
```{r}
fA_reduced <- 'log(response.time) ~ session * problem_order + problem_set + pain + physical_symptoms + discomfort + sensations + (1|participant_id)'
mA_reduced <- lmer(fA_reduced, data = data, REML = FALSE)
anova(m0, mA_reduced)
```

$p < 0.05$, so TMS condition is significant and we can move on to pinpointing the source by examining individual effects.

## TMS fixed effects

```{r}
# revisit the model summary
summary(m0)
```

The marginal fixed effect of TMS condition is not statistically significant ($t = 0.839$, $p = 0.4052$), and neither is the interaction of TMS condition with session number ($t = 0.069$, $p = 0.9451$).

However, the interaction of TMS condition with problem order is statistically significant ($t = -5.099$, $p = 3.54 \times 10^{-7}$), though small (estimate $= -0.0014618$). This implies a small multiplicative effect of TMS condition on response time that is dependent on the order in which problems are presented. Specifically, when in the control group for TMS condition, response times decrease as the test goes on, following a gradual exponential decay of $\times e^{-3.619 \times 10^{-4}} \approx 0.9996$ per problem (because the problem order coefficient is negative: estimate $= -3.619 \times 10^{-4}$). In contrast, when under the influence of TMS to the working memory region, response times decrease faster as the test goes on, following a steeper exponential decay of $\times e^{(-3.619 \times 10^{-4}) + (-1.462 \times 10^{-3})} \approx 0.9982$ per problem (because the interaction coefficient is negative: estimate $= -1.462 \times 10^{-3}$).

There is also a three-way interaction of TMS condition, session number, and problem order that is not significant at the 95% threshold, but is significant at the 90% threshold ($t = 1.902$, $p = 0.0573$, estimate $= 0.0007717$).

## TMS random effects

The selected model `m0` has a person-dependent TMS effect represented by a random intercept (`participant_id:TMS_Type`). Next, we check whether it is statistically significant using parametric bootstrapping.

```{r}
# Simple thread-safe append function
safe_write <- function(row, fout, lock_file) {
  lck <- lock(lock_file) # Acquire lock
  # on.exit(unlock(lck))  # Ensure unlock on finish or error

  sink(
    file = fout,
    append = TRUE,
    type = "output",
    split = FALSE
  )
  cat(row, " \n")
  sink()
  unlock(lck) # Release lock
}
```

```{r}
# Callback function for each bootstrap simulation
get_estimates <- function(m, save = FALSE, fout, effect_order, f_lock) {
  # Store variance estimates for all random effects
  random_effects <- as.data.frame(VarCorr(m)) %>%
    rename(Groups = grp, Name = var1, Variance = vcov, Std.Dev. = sdcor) %>%
    select(-var2)
  vv <- sapply(effect_order, function(eff) {
    (random_effects %>% filter(Groups == eff))$Variance
  }) # this is to ensure that they appear in the same order every time

  if (save) {
    if (missing(fout) || missing(effect_order)) {
      warning(
        paste0(
          "get_estimates failed to save: ",
          "fout or effect_order not provided"
        )
      )
    } else {
      safe_write(paste0(vv, collapse = ", "), fout, f_lock)
    }
  }

  return(vv)
}
```

```{r}
VarCorr(m0)
```

```{r bootmer_b0_2, cache=TRUE}
ordered_factors <- as.data.frame(VarCorr(m0))$grp
nsim <- 10000
# 10k sims took about 2.5 hrs

psve_log <- sprintf("%s/bootstrap_%i.csv", logtime_outputdir, nsim)
file.create(psve_log)
lock_file <- paste0(psve_log, ".lock") # Define a lock file
cat(psve_log, "\n")

# write the header
cat(paste0(ordered_factors, collapse = ","), "\n", file = psve_log)

b0 <- bootMer(
  m0,
  FUN = function(m) {
    get_estimates(
      m,
      save = TRUE,
      psve_log,
      ordered_factors,
      lock_file
    )
  },
  nsim = nsim,
  #.progress = "txt"
  parallel = "multicore",
  ncpus = 64
)
```

```{r}
b0_df <- data.frame(b0$t)
quantile(b0_df$participant_id.TMS_Type, c(0.025, 0.975))
```

The TMS random effect has a 95% confidence interval of $[0.001875637, 0.008148756]$, so is statistically significant. To interpret the effect size, we compute the PVE and its CI.

```{r}
estimates <- get_estimates(m0, effect_order = ordered_factors)
names(estimates) <- ordered_factors
pve_estimate <- estimates["participant_id:TMS_Type"] / sum(estimates)
cat(sprintf("Point estimate of PVE(participant_id:TMS_Type): %f", pve_estimate))

b0_df$pve_TMS <- b0_df$participant_id.TMS_Type / rowSums(b0_df)
quantile(b0_df$pve_TMS, c(0.025, 0.975))
```

So, the TMS random effect explains about 9.15% of the variance in log-transformed response times, with a 95% confidence interval of $[2.97\%, 13.08\%]$.

```{r}
estimates["participant_id:TMS_Type"] / (estimates["participant_id:TMS_Type"] + estimates["participant_id"])
```

It also explains about 30% of the structural variance among participants.
