---
title: "3-back Reaction Time Analysis"
output: html_document
date: "2025-1-29"
---

## Packages

```{r}
# import packages
library(lme4)     # mixed models
library(dplyr)    # dataframe manipulation
library(readr)    # reading in the data
library(emmeans)  # likelihood ratio testing
library(purrr)    # for map()
library(lattice)  # for xyplot()
library(TAF)      # setting up output file directory structure during model fitting
library(moments)  # skewness for selecting between additive and multiplicative models of response times



#If you get an error, you likely need to install the packages using the "install.package('insert-package-name-here')" command
```

## Notebook configurable settings

```{r}
# name the directories to read data from and write output to
input_version = 9
output_version = 'n-back'
nback_version = '3-back'

```

make an overall output directory (model summaries, stats, etc. will all go here)

```{r}
#setwd("..")
outputdir = sprintf("%s/../journal-candidate-models/regression-models-%s",getwd(), output_version)
```

```{r eval=FALSE}

mkdir(outputdir)
```

## Loading Data

```{r}
setwd("..")
input_file = sprintf("%s/preprocessed-data/full-data-version-%d/nback-full-data.csv", getwd(), input_version)
data <- read.csv(input_file)


#Transforming the session number such that it is treated as a character and not a value
data <- transform( 
  data, session = as.character(session)) 

head(data)
```

Removing irrelevant data from dataframe

```{r Nans}
#remove rows that have no values - nothing actually happens here. Everyone saw the stimuli
#data <- data[data$response.time != 0, ] 


#Removing the one participants that didn't listen to instructions:
participants_to_remove <- c("REDACTED_20")
data <- subset(data, !(participant_id %in% participants_to_remove))
```

Making sure we are only looking at one type of n-back at a time and only instance of correct response

```{r}
table(data$correct_answer)
```

```{r}
min(data$problem_order)
```

```{r}
data <- data[data$correct_answer == 'True', ] #only counting the onces they got correct


data <- data[data$nback_type == nback_version, ] #1-back , 2-back , or 3-back

```

#Determining what is the mean response.time

```{r}
mean_value <- mean((data$response.time))
print(mean_value)

```

Centralizing the response.time so that it is more normally distributed. Two-back is generally normally distributed as seen in the bottom graph so no transformation is required

```{r}
hist((data$response.time))
```

#Checking number of participants is n=27

```{r}

id <- unique(data$participant_id)
id #all unique IDs
length(id) #number of participants
```

#Checking distribution of of observartions per factor load in n-back

```{r}
count_table <- table(data$TMS_Type, data$nback_type)

# Print the result
print(count_table)
```

## Select independent variables

1.  `participant_id` (person) : random intercept
    -   People are drawn from an underlying distribution; we aren't interested in learning coefficients specific to the members of our sample, but rather in what they tell us about that underlying population. Random effects tell us about the distribution rather than the elements observed.

    -   This variable is used to model person-specific attributes, such as a "skill," translating to overall faster or slower response times, that can be different for each person
2.  `shown_letter` (letter presented) : random intercept
    -   The underlying distribution of stimuli is a little less clear than that of participants: specifically since stimuli in N-back are just a random fixed set of letters. However, prior work on N-back have found that some letters may be easier to remember than others. We control for that here.
3.  `TMS_Type` (TMS treatment condition, i.e. brain region stimulated) : fixed & random intercepts
    -   Based on previous literature, we hypothesize that the brain region stimulated may have an effect that is systematic (i.e. very consistent between people, topics, etc.), quantitatively heterogeneous (i.e. directionally consistent, but widely varying in magnitude), or even a qualitatively heterogeneous (i.e. varying in direction, possibly in conjunction with magnitude). Therefore, we assess the goodness of fit for candidate models containing `TMS_Type` as a fixed effect, random effect, both, or neither, in order to test this hypothesis.
4.  `session` (chronological numbering of the session, per participant) : fixed intercept
    -   This is used to account for learning/training effect, i.e. participants getting better at test-taking under our specific test instrument over the course of the study. For example, they may become more familiar with the question structure, or become better at recalling old material as they do more practice questions. We expect this to be systematic (uniformly decreasing response time), so we model it as a fixed effect.

## Specify candidate predictor structures for mixed model formulas

```{r}
fixed_effects = c(# no fixed effects
                  "", 
                  
                  # no interactions
                  "TMS_Type", 
                  #"nback_type", #does not make sense since we are looking at each n-back seperately
                  "session",
                  #"TMS_Type + nback_type", 
                  #"nback_type + session", 
                  #"TMS_Type + session", 
                  #"TMS_Type + nback_type + session",
                  "problem_order", #the letter that they saw at what time
                  
                   "pain",
                  "physical_symptoms",
                  "discomfort", 
                  "sensations",
                  
                  # 2-way interactions
                  #"TMS_Type * nback_type", 
                  #"nback_type * session", 
                  "TMS_Type * session", 
                  "TMS_Type * problem_order",
                  "session * problem_order",
                  "TMS_Type * session * problem_order",
                  
                   "TMS_Type * pain",
                  "TMS_Type * physical_symptoms",
                  "TMS_Type * discomfort",
                  "TMS_Type * sensations"
                  
                  
                  
                  # 1- and 2-way interactions
                  #"TMS_Type * nback_type + session", 
                  #"nback_type * session + TMS_Type", 
                  #"condition * session_num + domain", 
                  
                  # multiple 2-way interactions
                  #"TMS_Type * nback_type + nback_type * session" #,
                  #"condition * domain + condition * session_num",
                  #"domain * session_num + condition * session_num",
                  #"condition * domain + domain * session_num + condition * session_num",
                  # 3-way interactions
                  #"condition * domain * session_num"
                  )
```

```{r}
random_effects = c(#"",                                   # (1)
                   "participant_id", 
                   "shown_letter",                    
                   #"condition",                          # (2)
                   "participant_id:shown_letter",                   
                   #"participant_id:nback_type",          # We are only looking at one n-back type
                   "participant_id:TMS_Type",
                   "shown_letter:TMS_Type",          
                   "participant_id:shown_letter:TMS_Type"#,          
                   #"participant_id:nback_type:TMS_Type"  # We are only looking at one n-back type
                   )


```

In this first cell, we enumerate random effect structures by taking all combinations of the random effects specified.

```{r}
random_effects = map(random_effects, function (s) sprintf("(1|%s)", s))

random_exprs = c()
N_random = length(random_effects)
for (k in 1:N_random) {
  iii = combn(1:N_random, k, simplify=FALSE)
  for (ii in iii) {
    expr = paste(random_effects[ii], collapse = " + ")
    random_exprs = append(random_exprs, expr)
  }
}
```

Now, we combine the response variable, fixed effects, and random effects together to generate our candidate model structures.

```{r}
loglinear_formulae = c()
for (fe in fixed_effects) {
  for (re in random_exprs) {
    f = ""
    if (fe == "") {
      f = sprintf("(response.time) ~ %s", re)
    } else {
      f = sprintf("(response.time) ~ %s + %s", fe, re)
    }
    loglinear_formulae = append(loglinear_formulae, f)
  }
}

print(length(loglinear_formulae))
```

### Model fitting (\~2 minutes)

Here, we fit all the models. Of note, we use MLE rather than REML for compatibility with the AIC model selection strategy.

We also store BIC (an alternative model selection criterion) in addition to AIC here. The two have different theoretical properties, both of which may be useful. Both describe the log likelihood of a fitted model, with a penalty for model complexity. The penalty in BIC is higher than that for AIC, so optimal-BIC models tend to be smaller (with fewer parameters) than optimal-AIC models. We have done our analysis relying on AIC, but we provide BIC-based model selection here in case the reader is interested.

```{r eval=FALSE}
if (TRUE) {
  # make an output directory for the output of all the candidate linear models
  logtime_outputdir = 
    sprintf("%s/threeback_v9_time_model_summaries", outputdir)
  mkdir(logtime_outputdir)
  
  # make an output file for the AIC and BIC of each model
  metricsfile = sprintf("%s/metrics.csv", logtime_outputdir)
  file.create(metricsfile)
  
  sink(file = metricsfile,
       append = FALSE,
       type = "output",
       split = FALSE
       )
  
  cat("formula,AIC,BIC\n")
  
  sink()
  
  for (f in loglinear_formulae) {
    # make an output directory for the output of this specific candidate model
    f_outputdir = sprintf("%s/%s", logtime_outputdir, gsub(" ", "", f, fixed = TRUE))
    mkdir(f_outputdir)
    
    # make the individual output file
    summaryfile = sprintf("%s/summary.txt", f_outputdir)
    
    file.create(summaryfile)
    
    # print a delimiter for console output
    cat(sprintf("\n---------------------------------------------\n%s\n", f))
    
    # fit the model
    m = lmer(f, data=data, REML=FALSE)
    
    # send the model summary to the summary file
    sink(file = summaryfile,
         append = FALSE,
         type = "output",
         split = FALSE
         )
    
    print(summary(m))
    
    sink()
    
    # send the AIC and BIC to the metrics file
    metrics = sprintf("%s,%f,%f\n", f, AIC(m), BIC(m))
    
    sink(file = metricsfile,
         append = TRUE,
         type = "output",
         split = FALSE)
    
    cat(metrics)
    
    sink()
  }
}
```

### Model selection

```{r}
# read in the AIC/BIC data
fin = sprintf("%s/threeback_v9_time_model_summaries/metrics.csv", outputdir)
log_time_models_da = read_csv(fin)
```

```{r}
head(log_time_models_da[order(log_time_models_da$AIC),], n = 10)
```

```{r}
head(log_time_models_da[order(log_time_models_da$BIC),], n = 10)
```

### Optimal AIC

```{r}
fA = log_time_models_da[order(log_time_models_da$AIC),]$formula[[1]]
cat(fA)
```

AIC favors '(response.time) \~ session \* problem_order + (1\|participant_id) + (1\|participant_id:TMS_Type)'

```{r}
fB = log_time_models_da[order(log_time_models_da$BIC),]$formula[[1]]
cat(fB)
```

BIC favors a different model so we should check what is best for data fitting

```{r}
m1 = lmer(fA, data=data, REML=FALSE)
summary(m1)
```

```{r}
m0 = lmer(fB, data=data, REML=FALSE)
summary(m0)
```

## Which model is better AIC or BIC selected

If Pr(\>Chisq) \< 0.05 → m1 is significantly better than m0 ✅ Keep m1, since the extra parameter improves model fit. If Pr(\>Chisq) \> 0.05 → No significant difference ❌ Prefer m0 (simpler model), since m1 doesn’t improve fit enough. If AIC is lower in m1 but p \> 0.05 If the AIC difference is ≥ 2, consider m1. If ΔAIC \< 2, prefer the simpler m0.

```{r}
anova(m1,m0)

```

We choose m1: (response.time) \~ session \* problem_order + (1 \| participant_id) + (1 \| participant_id:TMS_Type) because p \< 0.05

## CI

```{r}
if (TRUE) {
  pp0 = profile(m1)
}
```

```{r}
confint(pp0)
```

All random effects have CI that don't inlcude zero.

### Point estimate PVE

```{r}
get_pve = function(m, effect) {
  random_effects = as.data.frame(VarCorr(m)) %>% 
    rename(Groups = grp, Name = var1, Variance = vcov, Std.Dev. = sdcor) %>% 
    select(-var2)
  
  num = (random_effects %>% filter(Groups == effect))$Variance
  
  denom = sum(random_effects$Variance)
  
  return (c(num / denom))
}


get_pve(m1, "participant_id:TMS_Type")
```

The person-condition random effect explains 6.8% of the variance in the data, according to the point estimates.

## PVE confidence interval (\~5 minutes)

```{r eval=FALSE}
if (TRUE) {
  b0 = bootMer(m0, 
               FUN = function (m) {return(get_pve(m, "participant_id:TMS_Type"))}, 
               nsim = 1000,
               .progress = "txt")
}
```

```{r eval= FALSE}
quantile(b0$t, c(0.025, 0.975))
```
